---
title: "individual2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
defense <- read_excel('cyclonesFootball2020.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2020.xlsx', sheet='Offensive')
biography <- read_excel('cyclonesFootball2020.xlsx', sheet='Biography')
```


# setup caitlin did


## Part One
## Question One
```{r}
defense <- rename(defense, 'Opponent'='Opponent_Opponent')
defense$Opponent <- as.factor(defense$Opponent)
defense$Name <- as.factor(defense$Name)

offense <- rename(offense, 'Opponent'='Opponent_Opponent')
offense$Opponent <- as.factor(offense$Opponent)
offense$Name <- as.factor(offense$Name)

biography$Name <- as.factor(biography$Name)
```

## Question 2
```{r}
defense <- defense %>%
  mutate(across(Tackles_Solo:Pass_PB, as.numeric))
str(defense)

offense <- offense %>%
  mutate(across(Receiving_REC:Passing_INT, as.numeric))
str(offense)


biography <- separate(biography, Height, c('Feet', 'Inches'), sep='-')
biography <- biography %>%
  mutate(across(Feet:Weight, as.numeric))
head(biography)

defClean <- defense
offClean <- offense
bioClean <- biography
```

# Part Two
### 1

```{r}
offCleanLonger <- offClean %>%
  pivot_longer(Receiving_REC:Passing_INT,names_to = 'stat',values_to = 'value' )

offCleanest <- drop_na(offCleanLonger)
```


### 2

```{r}
offCleanest$stat <- as.factor(offCleanest$stat)
ggplot(offCleanest, aes(x=value))+geom_histogram(bins = 15)+facet_wrap(~ stat, scales = 'free')
```

Passing_INT is skewed to the right with a mean at 0 passing interceptions. There is at least one outlier at 3 passing interceptions in one game.
Passing_TD has a mean of 0 per game and as the number of passing touchdowns tends to decrease as the number increases.
There are more instances of players with 0 passing yards per game than any other category, but we can see that there are likely a few dominant players that average between 200 and 300 yards per game.
Receiving_REC tends to decrease as the number of receptions increases. There seem to be a couple outliers at around 10 receptions a game.
Offensive players had either 0 or 1 receiving touchdown per game, with 0 touchdowns per game being more common.
Receiving_YDS decreases as the number of years increases, with potential outliers at around 110 yards and 120 yards. The median value was 0 yards.
Rushing_ATT tends to decrease as the amount increases, but the most common values for Rushing_ATT were 0 and around 7-8.There seem to be a couple outliers at around 30.
Rushing_TD decreases at the number of touchdowns increases.The most common value was 0.
Rushing_YDS decreases asthe number of yards increases and there seems to be potential outliers closer to 200 yards.
All distributions are skewed to the right and have outliers at the extreme right tail values.

### 3
```{r}
offCleanWide<- offCleanest %>%
  filter(stat == "Receiving_YDS" & ( Opponent == "Oklahoma" | Opponent == "Oregon")) %>%
  pivot_wider(names_from='Opponent',values_from="value", values_fn=mean) 
 
head(offCleanWide)

ggplot(offCleanWide, aes(x=Oklahoma, y= Oregon))+geom_point()+xlim(0,100)+ylim(0,100)

```
We can see that the defense was better against Oklahoma because every player had higher Recieving Yards against Oklahoma than against Oregon. THis is visible because all of the points are below the x=y line. 



